Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference

### 2025-01-22 src/tex/ms.tex
[showyourwork] overleaf sync

For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with 1000 points. This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars. Our method takes around $10$ seconds to build a posterior for all six parameters for a dataset size of 1000 stars (each time sampling the single star posterior with 1000 points and fitting for the Gaussian parameters). Our combined runtime for the NPE training plus sampling is hence 610 seconds for 1000 stars hence our methods making it more than $1.180$ times faster than current HMC methods \citep[cf.][who needs $40$h for only 200 stars]{Philcox_2019}. Hence, shorter computing times make it feasible to use orders of magnitudes more observations.

>>>>

For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with 1000 points. This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars. Our method takes around $10$ seconds to build a posterior for all six parameters for a dataset size of 1000 stars (each time sampling the single star posterior with 1000 points and fitting for the Gaussian parameters). Our combined runtime for the NPE training plus sampling is hence 610 seconds for 1000 stars hence our methods making it more than $1.180$ times faster than current HMC methods \citep[cf.][who needs $40$h for only 200 stars]{Philcox_2019}. But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim72.000$. \textcolor{red}{check if those speed-up values are correctly calculated...}
Hence, shorter computing times make it feasible to use orders of magnitudes more observations.
